---
title: "Applied Predictive Modelling C1"
date: "Tuesday, May 26, 2015"
output: html_document
---
```{r, message=FALSE}
library(AppliedPredictiveModeling); library(caret); library(corrplot); library(e1071); library(lattice); library(dplyr)
```

##Computing

### Selection

```{r}
# apropos() function will search any loaded R function for given term
apropos('confusion')

# to find function in any packahe, RSiteSearch() is handy

#RSiteSearch('confusion', restrict = 'functions')
```

The segemenatation data nice example set- lets play with subsetting

```{r}
data(segmentationOriginal)

# data already split into train and test
segData <- filter(segmentationOriginal, Case == 'Train')
```

The Class and Cell fields can be saved into seperate vectors, then removed from the main object

```{r}
cellID <- segData$Cell
class <- segData$Class
case <- segData$Case
# remove columns
segData <- segData[, -(1:3)]
```
Want to remove column names with 'Status', as these are binary versions of predictors

```{r}
statusColNum <- grep('Status', names(segData))
segData <- segData[, -statusColNum]
```

### Transformation

Some of the data are significantly skewed

```{r}
skewness(segData$AngleCh1)
# Can apply skewness over all columns 
skewValues <- apply(segData, 2, skewness)
head(skewValues)
```

Can use this as first hint- investigate skewbed predictors further.

MASS package contains boxcox() function. This function estimates \( \lambda \) lambda. Use caret function BoxCoxTrans() to apply the transformation.

```{r}
Ch1AreaTrans <- BoxCoxTrans(segData$AreaCh1)
Ch1AreaTrans

#origional data
head(segData$AreaCh1)

#after transform
predict(Ch1AreaTrans, head(segData$AreaCh1))
```
The caret function preProcess() applies transformations to set of predictors. This can include a bunch or preprocessings (see example in a bit)

prcomp() function in base R:

```{r}
pcaObject <- prcomp(segData, center=TRUE, scale. = TRUE)
#calculate percentage variance of components
percentVariance <- pcaObject$sd^2 / sum(pcaObject$sd^2) * 100
percentVariance[1:3]
```

Transformed values are stored in pcaObject as sub-object x:

```{r}
head(pcaObject$x[, 1:5])
```

The sub-object rotation stores the variable loadings

```{r}
head(pcaObject$rotation[, 1:3])
```

got to page 54 - spatialSign()
