---
title: "Applied Predictive Modelling C1"
date: "Tuesday, May 26, 2015"
output: html_document
---
```{r, message=FALSE}
library(AppliedPredictiveModeling); library(caret); library(corrplot); library(e1071); library(lattice); library(dplyr)
```

##Computing

### Selection

```{r}
# apropos() function will search any loaded R function for given term
apropos('confusion')

# to find function in any packahe, RSiteSearch() is handy

#RSiteSearch('confusion', restrict = 'functions')
```

The segemenatation data nice example set- lets play with subsetting

```{r}
data(segmentationOriginal)

# data already split into train and test
segData <- filter(segmentationOriginal, Case == 'Train')
```

The Class and Cell fields can be saved into seperate vectors, then removed from the main object

```{r}
cellID <- segData$Cell
class <- segData$Class
case <- segData$Case
# remove columns
segData <- segData[, -(1:3)]
```
Want to remove column names with 'Status', as these are binary versions of predictors

```{r}
statusColNum <- grep('Status', names(segData))
segData <- segData[, -statusColNum]
```

### Transformation

Some of the data are significantly skewed

```{r}
skewness(segData$AngleCh1)
# Can apply skewness over all columns 
skewValues <- apply(segData, 2, skewness)
head(skewValues)
```

Can use this as first hint- investigate skewbed predictors further.

MASS package contains boxcox() function. This function estimates \( \lambda \) lambda. Use caret function BoxCoxTrans() to apply the transformation.

```{r}
Ch1AreaTrans <- BoxCoxTrans(segData$AreaCh1)
Ch1AreaTrans

#origional data
head(segData$AreaCh1)

#after transform
predict(Ch1AreaTrans, head(segData$AreaCh1))
```
The caret function preProcess() applies transformations to set of predictors. This can include a bunch or preprocessings (see example in a bit)

prcomp() function in base R:

```{r}
pcaObject <- prcomp(segData, center=TRUE, scale. = TRUE)
#calculate percentage variance of components
percentVariance <- pcaObject$sd^2 / sum(pcaObject$sd^2) * 100
percentVariance[1:3]
```

Transformed values are stored in pcaObject as sub-object x:

```{r}
head(pcaObject$x[, 1:5])
```

The sub-object rotation stores the variable loadings

```{r}
head(pcaObject$rotation[, 1:3])
```

For spatial sign transforms REMEMBER to center and scale first. Spatial sign for models which are sensitive to outliers

```{r}
#spatObject <- spatialSign(segData)
```

The segData does not have any missing values, but for data with these can use impute package with knn imputation impute.knn()

The preProcess() function in caret is handy to apply all the preprocessing

```{r}
trans <- preProcess(segData,
                 method = c("BoxCox", "center", "scale", "pca"))
trans

#apply transformations
tranformed <- predict(trans, segData)
```
The order of preprocessing is center, scale, inputation, feature extraction then spatial sign.

### Filtering

Use nearZeroVar() to find predictors with near zero variance

```{r}
nearZeroVar(segData)
```

For correlations use cor()

```{r, message=FALSE}
correlation <- cor(segData)
correlation[1:4, 1:4]
corrplot(correlation, order = "hclust")
```

To filter based on correlations, use findCorrelation()

```{r}
highCorr <- findCorrelation(correlation, cutoff = .75)
filteredSegData <- segData[, -highCorr]
```
### Dummies

Use dummies for tree based models- makes splits more interpretable!!

```{r}
library(caret)
data(cars)
#seems cars are already in dummy form
#carSub <- select(cars, Price, Mileage, Type)
#head(cars)
#levels(carSub$Type)

#simpleMod <- dummyVars(~Mileage + Type,
#                       data = carSubset,
#                       levelsOnly = TRUE)

# makes Type a dummy now
# USE predict() method on dataset in conjunction with dummyVars

#predict(simpleMod, head(carSubset))

#if you want an interaction term, use :

#withInteraction <- dummyVars(~Mileage + Type + Mileage:Type,
#                    data = carSubset, levelsOnly = TRUE)

```


